Teaching Low-Level Concurrency with \pebbles and \landslide

### 2016-08-27 sigcse-17/eval.tex
sigcse: save some frames^Wlines and add a conclusion paragraph

\begin{table}[t]
	\begin{tabular}{r|cc|cc}
	%det bugs found	det bugs fixed	races fixed	races found
	& \multicolumn{4}{c}{Number of groups} \\
	Number & \multicolumn{2}{c|}{Deterministic bugs} & \multicolumn{2}{c}{Nondeterministic bugs} \\
	of bugs	& found & fixed & found & fixed \\
	\hline
	0	& 27	& 32	& 15	& 23	\\
	1	& 6	& 5	& 7	& 11	\\
	2	& 9	& 6	& 12	& 8	\\
	3	& 2	& 1	& 7	& 2	\\
	4	& 1	& 1	& 3	& 1	\\
	5	& 2	& 2	& 2	& 1	\\
	11	& 	& 	& 1	& 1	\\
	\hline
	%Total groups
	%	& 47	& 47	& 47	& 47	\\
	Total bugs
		& 44	& 34	& 85	& 53	\\
	\end{tabular}
	\caption{Summary of how many bugs were found and/or fixed by how many groups.
	Each column counts how many groups found/fixed the number of bugs in each row;
	for instance, two groups found 5 concurrency bugs, one of which fixed all 5.}
	\label{tab:this-table-sucks-but-it's-the-best-i-got}
\end{table}

@@

\begin{figure}[t]
	\includegraphics[width=0.48\textwidth]{p2-distribution-logos.pdf}
	\caption{Distribution of thread library grades.}
	\label{fig:p2-distribution}
\end{figure}
We also collected the overall thread library grades to evaluate whether \landslide helps students achieve more during the project.
Figure~\ref{fig:p2-distribution} shows the distribution of these grades.
Compared to students from the same semesters who opted not to use \landslide,
those who did scored on average 3\% higher. % that's rough, buddy.
%
%However, to account for the possible bias for opting-out students being more at the bottom of the class
%on account of not enough free time to volunteer for \landslide,
However, the latter distribution is potentially biased towards struggling students
who lack enough free time to volunteer for \landslide.
To account for this, we also checked grades from prior semesters when \landslide was not available.
%This distribution is indistiguishable from the experimental group,
Although the minimum grade among the experimental group is 17\% higher than the minimum from prior semesters,
the middle 50\% of the distributions are quite similar,
suggesting that \landslide's benefit is largely offset by the time investment required.
%
In future work, we will streamline the user experience to help students fix as many concurrency bugs with a smaller time investment.

>>>>

\begin{table}[t]
	\begin{center}
	\begin{tabular}{r|cc|cc}
	%det bugs found	det bugs fixed	races fixed	races found
	& \multicolumn{4}{c}{Number of groups} \\
	Number & \multicolumn{2}{c|}{Deterministic bugs} & \multicolumn{2}{c}{Concurrency bugs} \\
	of bugs	& found & fixed & found & fixed \\
	\hline
	0	& 27	& 32	& 15	& 23	\\
	1	& 6	& 5	& 7	& 11	\\
	2	& 9	& 6	& 12	& 8	\\
	3	& 2	& 1	& 7	& 2	\\
	4	& 1	& 1	& 3	& 1	\\
	5	& 2	& 2	& 2	& 1	\\
	11	& 	& 	& 1	& 1	\\
	\hline
	%Total groups
	%	& 47	& 47	& 47	& 47	\\
	Total bugs
		& 44	& 34	& 85	& 53	\\
	\end{tabular}
	\end{center}
	\caption{Summary of how many bugs were found and/or fixed by how many groups.
	Each column counts how many groups found/fixed the number of bugs in each row;
	for instance, two groups found 5 concurrency bugs, one of which fixed all 5.}
	\label{tab:this-table-sucks-but-it's-the-best-i-got}
\end{table}

@@

\begin{figure}[t]
	\includegraphics[width=0.48\textwidth]{p2-distribution-logos.pdf}
	\caption{Distribution of thread library grades.}
	\label{fig:p2-distribution}
\end{figure}
We also collected the overall thread library grades to evaluate whether \landslide helps students achieve more during the project.
Figure~\ref{fig:p2-distribution} shows the distribution of these grades.
Compared to students from the same semesters who opted not to use \landslide,
those who did scored on average 3\% higher. % that's rough, buddy.
%
%However, to account for the possible bias for opting-out students being more at the bottom of the class
%on account of not enough free time to volunteer for \landslide,
However, the latter distribution is potentially biased towards struggling students
who lack enough free time to volunteer for \landslide.
%To account for this,
Thus, we also checked grades from prior semesters when \landslide was not available.
%This distribution is indistiguishable from the experimental group,
Although the minimum grade among the experimental group is 17\% higher than the minimum from prior semesters,
the middle 50\% of the distributions are quite similar,
suggesting that \landslide's measurable benefit is largely offset by the time investment required.
%
%In future work, we will streamline the user experience to help students fix as many concurrency bugs with a smaller time investment.

### 2016-08-27 sigcse-17/paper.tex
sigcse: save some frames^Wlines and add a conclusion paragraph

We aspire to expand the scope of these projects in response to
typical hardware platforms'
increasing complexity,
especially massive multi-core.
To that end, we wish to reduce the time students spend
debugging concurrency bugs.
We present \landslide, a
%systematic testing
stateless model checker which
offers a more principled approach than stress testing.
% edupar version
%We report on our preliminary experience working with students using \landslide to find previously-overlooked bugs in their own code, and discuss future methods for making \landslide more accessible to struggling students.
% sigcse version
We report on our experience giving students \landslide during the thread library
%and kernel projects.
project,
%Although kernel-mode \landslide requires manual instrumentation, testing thread libraries is fully automatic.
showing that it allows them to find and fix previously-overlooked bugs.
%% no :( :(
% Further, %even struggling students can take advantage of it,
% \landslide's users achieve higher concurrency scores on average in subsequent projects,
% suggesting that debugging with a model checker improves student retention of concurrency principles.

@@

\ccsdesc[500]{Software and its engineering~Software maintenance tools}
\ccsdesc[300]{Computing methodologies~Concurrent algorithms}

@@

We are presently extending \landslide to support Pintos kernels \cite{pintos},
%used at many top CS schools including Stanford, Berkeley, and University of Chica
another instructional kernel popular at many top CS schools.
%a more widely-used instructional kernel than \pebbles.
Because Pintos mandates more internal APIs than \pebbles,
\landslide can more easily instrument student code without requiring manual annotations.
In a preliminary experiment, we tested 79 student Pintos implementations with \landslide,
finding 59 concurrency bugs with 10 CPU-hours allocated to each of 3 test cases \cite{quicksand-anonymized}.
In future work, we hope to fully automate Pintos instrumentation, and offer \landslide at other schools beyond \cmu.

%\fourten, the undergraduate operating systems course at \cmu, teaches concurrency in a

>>>>

We aspire to expand the scope of these projects in response to
typical hardware platforms'
increasing complexity,
especially massive multi-core.
To that end, we wish to reduce the time students spend
debugging concurrency bugs.
We present \landslide, a
%systematic testing
stateless model checker which
offers a more principled approach than stress testing.
% edupar version
%We report on our preliminary experience working with students using \landslide to find previously-overlooked bugs in their own code, and discuss future methods for making \landslide more accessible to struggling students.
% sigcse version
We report on our experience giving students \landslide during the thread library
%and kernel projects.
project,
%Although kernel-mode \landslide requires manual instrumentation, testing thread libraries is fully automatic.
showing that it allows them to find and fix many previously-overlooked concurrency bugs.
%% no :( :(
% Further, %even struggling students can take advantage of it,
% \landslide's users achieve higher concurrency scores on average in subsequent projects,
% suggesting that debugging with a model checker improves student retention of concurrency principles.

@@

\ccsdesc[500]{Software and its engineering~Software maintenance tools}
%\ccsdesc[300]{Computing methodologies~Concurrent algorithms}

@@

\fourten, the undergraduate operating systems course at \cmu,
features low-level thread library and kernel projects
in which students tackle challenging concurrency problems.
We have introduced \landslide, a stateless model checker,
as an alternative debugger to conventional stress testing.
In our evaluation over three semesters,
\landslide helped more than half of volunteering students find and fix previously-overlooked concurrency bugs.
So far, however, this benefit is not reflected in the grades students ultimately achieved.
In future work, we will streamline the user experience to help students fix as many concurrency bugs with a smaller time investment.
%helping students achieve more overall.

We are presently extending \landslide to support Pintos kernels \cite{pintos},
%used at many top CS schools including Stanford, Berkeley, and University of Chica
another instructional kernel popular at many top CS schools.
%a more widely-used instructional kernel than \pebbles.
Because Pintos mandates more internal APIs than \pebbles,
\landslide can easily instrument student code without requiring manual annotations.
In a preliminary experiment, we tested 79 student Pintos implementations with \landslide,
finding 59 concurrency bugs with 10 CPU-hours allocated to each of 3 test cases \cite{quicksand-anonymized}.
In future work, we hope to fully automate Pintos instrumentation, and offer \landslide at other schools beyond \cmu.
